Introducing statistics with \\texttt{intRo}

### 2014-11-20 student-paper-competition-2014/paper.Rmd
more estensibility
worked on reactive programming

Beyong being simple, \texttt{intRo} is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. In figure \ref{fig:ui}, we highlight the five elements that comprise the \texttt{intRo} interface.

@@

## Extensible
\label{sec:extensible}
On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules. This is true of the back end as well. Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time. What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application.

@@

>>>>

Beyond being simple, \texttt{intRo} is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. In figure \ref{fig:ui}, we highlight the five elements that comprise the \texttt{intRo} interface.

@@

## Extensible
\label{sec:extensible}
On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules. This is true of the back end as well. Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time. What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application. This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.

@@

Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@reactive2013] As implemented by \texttt{shiny} [@shiny], results automatically update when users interact with the interface.

\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, this carries throughout the entire session. In a typical \texttt{shiny} app, users interact with inputs, which in turn yield different results. Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application. This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.

In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability. At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability. For a tangible example, consider the graphical summaries module. We only allow variables of a type consistent with the selected plot to be displayed. This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application. On the flip side of this, \texttt{shiny} allows much higher flexibility. For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}. [Maybe add a figure here?] However, \texttt{shiny} does have limits in its flexibility based on the designers decisions for usability. One current example the slider element. This element allows for fixed width steps from its minimum to its maximum. The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they much be written in plain JavaScript. This is an example of a decision made by the developers of \texttt{shiny} to limit functionality in favor of usability of their package.
